\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces}

### 2020-03-02 submission/manuscript.tex
OTU taxonomy file added to .gitignore.

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of eukaryotic,
chloroplast, mitochondrial and no relative sequences
(\autoref{supp-nseq_notus}). A total of 35 samples originating from
epiphytic archaeal and bacterial communities associated with surfaces of
the seagrass \emph{Cymodocea nodosa} and macroalga \emph{Caulerpa
cylindracea} were analyzed. In addition, 18 samples (one of the samples
was sequenced two times) originating from picoplankton archaeal and
bacterial communities in the surrounding seawater were also processed.
The number of reads per sample ranged between 8,407 and 77,465 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off that is a common
observation in high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences mentioned before reads were clustered into 28,726 different
OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
normalized to the minimum number of sequences, 8,407
(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,007
different OTUs that ranged from 366 to 1,998 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes of richness
and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
Shannon (Jost, 2006) and Inverse Simpson were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
150.8 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
communities in the surrounding seawater (Number of OTUs, 528.1 ± 135.8
OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
large dissimilarities. Seawater communities richness was stable during
the studied period with the exception of one sampling point in December
2017 when larger values were observed. \emph{Cymodocea nodosa}
communities showed a slow increase towards the end of the study, while
\emph{Caulerpa cylindracea} (invaded and noninvaded) communities were
characterized by slightly larger values in Spring and Summer in
comparison to Autumn and Winter (\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient were, respectively, calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from a
particular environment. The highest proportion of shared OTUs and
community was found between invaded and noninvaded \emph{Caulerpa
cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
values were calculated between seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{Cymodocea nodosa}
and \emph{Caulerpa cylindracea} were approximately in the middle between
these two extremes. To assess seasonal changes in the proportion of
shared OTUs and communities the Jaccard's and Bray-Curtis Similarity
Coefficients were calculated between consecutive sampling points
(\autoref{shared}). Both coefficients showed similar trends. Temporal
proportional changes were more pronounced for seawater in comparison to
\emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
associated communities (\autoref{shared}). To further disentangle the
environmental and seasonal community dissimilarity a Principal
Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU
abundances was applied. It showed a clear separation between planktonic
and surface associated communities (\autoref{pcoa}). In addition, a
separation of epiphytic bacterial and archaeal communities based on host
species was determined. This separation was further supported by ANOSIM
(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of seawater
communities indicated a separation between Spring, Summer and
Autumn/Winter samples (ANOSIM, R = 0.64, \emph{p} \textless{} 0.001).
Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
showed a similar pattern (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01),
while communities from the surfaces of \emph{Caulerpa cylindracea}
indicated a non so strongly supported, as in previous cases, separation
between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31,
\emph{p} \textless{} 0.05) (\autoref{pcoa}).

@@

\leavevmode\hypertarget{ref-Armstrong2000}{}%
Armstrong, E., Rogerson, A., and Leftley, J.W. (2000) The Abundance of
Heterotrophic Protists Associated with Intertidal Seaweeds.
\emph{Estuarine, Coastal and Shelf Science} \textbf{50}: 415--424.

>>>>

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of eukaryotic,
chloroplast, mitochondrial and no relative sequences
(\autoref{supp-nseq_notus}). A total of 35 samples originating from
epiphytic archaeal and bacterial communities associated with surfaces of
the seagrass \emph{Cymodocea nodosa} and macroalga \emph{Caulerpa
cylindracea} were analyzed. In addition, 18 samples (one of the samples
was sequenced two times) originating from picoplankton archaeal and
bacterial communities in the surrounding seawater were also processed.
The number of reads per sample ranged between 8,407 and 77,465 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off that is a common
observation in high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences mentioned before reads were clustered into 28,726 different
OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
normalized to the minimum number of sequences, 8,407
(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,020
different OTUs that ranged from 346 to 2,010 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes of richness
and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
Shannon (Jost, 2006) and Inverse Simpson were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
1,693.5 ± 122.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,745.5 ±
150.9 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
OTUs, 1,056.7 ± 212.5 OTUs) and lower values for picoplankton
communities in the surrounding seawater (Number of OTUs, 531.7 ± 142.7
OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
large dissimilarities. Seawater communities richness was stable during
the studied period with the exception of one sampling point in December
2017 when larger values were observed. \emph{Cymodocea nodosa}
communities showed a slow increase towards the end of the study, while
\emph{Caulerpa cylindracea} (invaded and noninvaded) communities were
characterized by slightly larger values in Spring and Summer in
comparison to Autumn and Winter (\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient were, respectively, calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from a
particular environment. The highest proportion of shared OTUs and
community was found between invaded and noninvaded \emph{Caulerpa
cylindracea} (Jaccard, 0.36; Bray-Curtis, 0.77), while lower shared
values were calculated between seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{Cymodocea nodosa}
and \emph{Caulerpa cylindracea} were approximately in the middle between
these two extremes. To assess seasonal changes in the proportion of
shared OTUs and communities the Jaccard's and Bray-Curtis Similarity
Coefficients were calculated between consecutive sampling points
(\autoref{shared}). Both coefficients showed similar trends. Temporal
proportional changes were more pronounced for seawater in comparison to
\emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
associated communities (\autoref{shared}). To further disentangle the
environmental and seasonal community dissimilarity a Principal
Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU
abundances was applied. It showed a clear separation between planktonic
and surface associated communities (\autoref{pcoa}). In addition, a
separation of epiphytic bacterial and archaeal communities based on host
species was determined. This separation was further supported by ANOSIM
(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of seawater
communities indicated a separation between Spring, Summer and
Autumn/Winter samples (ANOSIM, R = 0.63, \emph{p} \textless{} 0.001).
Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
showed a similar pattern (ANOSIM, R = 0.54, \emph{p} \textless{} 0.001),
while communities from the surfaces of \emph{Caulerpa cylindracea}
indicated a non so strongly supported, as in previous cases, separation
between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31,
\emph{p} \textless{} 0.05) (\autoref{pcoa}).

@@

\leavevmode\hypertarget{ref-Armstrong2000}{}%
Armstrong, E., Rogerson, A., and Leftley, J. (2000) The Abundance of
Heterotrophic Protists Associated with Intertidal Seaweeds.
\emph{Estuarine, Coastal and Shelf Science} \textbf{50}: 415--424.
